A Simple Type Theory for Category Theory

### 2022-05-19 paper/paper.tex
some changes to the semantics section

\begin{definition}
  A strict double category is a monoid in the double category $Span(Cat)$.
\end{definition}

@@

\begin{definition}
  A DCD is an Equipment with displays when
  \begin{enumerate}
  \item The functor $(s,t) : C_d \to C_0$ is a split fibration, with
    restriction operation written as $(f,g)^*R$.
  \item The profunctor $C_\vdash$ respects restrictions in that
    \[ C_\vdash^{f',g'}(\Phi, (f,g)^*R) \cong C_\vdash^{f\circ f',g\circ g'}(\Phi, R) \]
  \end{enumerate}
\end{definition}
Specializing this property give us that $C_\vdash^{f,g}(\Phi,R) \cong
C_\vdash^{\id,\id}(\Phi,(f,g)^*R)$, which justifies that we only need
to give syntax for terms $\Phi \vdash t : R$ where $\Phi$ and $R$ have
the same covariant/contravariant dependencies.

@@

\begin{definition}
  A Category with Families consists of
  \begin{enumerate}
  \item A category $\textrm{Ctx}$
  \item with a terminal object $1 \in \textrm{Ctx}$
  \item A presheaf $\textrm{Ty} : \Set^\textrm{Ctx}$ of types in context.
  \item A presheaf $\textrm{Tm} : \Set^{\int \textrm{Tm}}$ of terms of
    a given type.
  \item With a context extension functor $\textrm{ext} : \int
    \textrm{Ty} \to \textrm{Ctx}$ with universal property:
    \[ \textrm{Ctx}(\Delta,\textrm{ext}(A)) \cong (\gamma : \textrm{Ctx}(\Delta,\textrm{ctx}(A))) \times \textrm{Tm}(A[\gamma])  \]
  \end{enumerate}
\end{definition}

@@

\begin{definition}
  Let $(\textrm{Ctx},\textrm{Ty},\textrm{Tm},\textrm{ext})$ be a
  category with families and let $\mathcal V$ be a DCD internal to
  $\Set^\textrm{Ctx}$. Then
  \begin{enumerate}
  \item $\mathcal V$ objects are representable as types if there is a
    type given by a section $\textrm{Ob} \in \textrm{Ty}(1)$ with a
    natural isomorphism
    \[ \textrm{Tm}(!_\Gamma^*\textrm{Ob}) \cong \textrm{Ob}(\mathcal V)_\Gamma \]
  \item $\mathcal V$ vertical arrows are representable as types if
    there is a type
    \[ \textrm{Tm}(((\gamma,\lceil a\rceil),\lceil b\rceil)^*\textrm{Functor} \cong
    \textrm{VArr}(\mathcal V_{\Delta})(a,b)
    \]
    and similarly horizontal displays are representable as types if
    \[ \textrm{Tm}(((\gamma,\lceil a\rceil),\lceil b\rceil)^*\textrm{Profunctor} \cong
    \textrm{Disp}(\mathcal V_{\Delta})(a,b)
    \]
  \item Closed $\mathcal V$ transformations are representable as types if
    \[ \textrm{Tm}() \]
  \end{enumerate}
\end{definition}

>>>>

\begin{definition}
  A strict double category is a monoid in the bicategory $Span(Cat)$.
\end{definition}

@@

\begin{definition}
  A DCD is an Equipment with displays when
  \begin{enumerate}
  \item The functor $(s,t) : C_d \to C_0 \times C_0$ is a split
    fibration, with restriction operation written as $(f,g)^*R$.
  \item The profunctor $C_\vdash$ respects restrictions in that
    \[ C_\vdash^{f',g'}(\Phi, (f,g)^*R) \cong C_\vdash^{f\circ f',g\circ g'}(\Phi, R) \]
  \end{enumerate}
\end{definition}
Specializing this property give us that $C_\vdash^{f,g}(\Phi,R) \cong
C_\vdash^{\id,\id}(\Phi,(f,g)^*R)$, which justifies that we only need
to give syntax for terms $\Phi \vdash t : R$ where $\Phi$ and $R$ have
the same covariant/contravariant dependencies.

@@

\begin{definition}
  A Category with Families consists of
  \begin{enumerate}
  \item A category $\textrm{Ctx}$
  \item with a terminal object $1 : \textrm{Ctx}$
  \item A presheaf $\textrm{Ty} : \widehat {\textrm{Ctx}}$ of types in context.
  \item A presheaf $\textrm{Tm} : \widehat{\int\textrm{Ty}}$ of terms of
    a given type.
  \item With a context extension functor $\textrm{ext} : \int
    \textrm{Ty} \to \textrm{Ctx}$ with universal property:
    \[ \textrm{Ctx}(\Delta,\textrm{ext}(A)) \cong (\gamma : \textrm{Ctx}(\Delta,\textrm{ctx}(A))) \times \textrm{Tm}(A[\gamma])  \]
  \end{enumerate}
\end{definition}

@@

\begin{definition}
  Let $(\textrm{Ctx},\textrm{Ty},\textrm{Tm},\textrm{ext})$ be a
  category with families and let $\mathcal V = (C,C_d)$ be a DCD internal to
  $\Set^\textrm{Ctx}$. Then
  \begin{enumerate}
  \item $\mathcal V$ objects are representable as types if there is a
    type given by a section $\textrm{Ob} \in \textrm{Ty}(1)$ with a
    natural isomorphism
    \[ \textrm{Tm}(!_\Gamma^*\textrm{Ob}) \cong C_{0}(\Gamma) \]
  \item $\mathcal V$ vertical arrows are representable as types if
    there is a type
    \[ \textrm{Tm}(Functor)(\Gamma)(a,b) \cong C_0(\Gamma)(a,b)
    \]
    and similarly horizontal displays are representable as types if
    \[ \textrm{Tm}(Prof)(\Gamma)(a,b) \cong \{ R \in C_d(\Gamma) \,|\, s(R) = a \wedge t(R) = b \}
    \]
  \item Closed $\mathcal V$ transformations are representable as types if
    \[ \textrm{Tm}(\textrm{End})(\Gamma)(a,R) \cong C_\vdash^{\id,\id}(r(a),R) \]
  \end{enumerate}
\end{definition}
